---
title: "Lab 5 - Spatial Dependence"
author: "ECOL 620 - Applications in Landscape Ecology"
date: "Spring 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This lab assignment highlights the use of a number of R packages to explore differing techniques for assessing spatial dependence. Download Laboratory #5 materials and store files locally on your computer. Read “Fletcher_2018_CH5” and open “lab_5_practice.Rmd” in R. 

```{r}
#load packages
library(pgirmess)         #for simple correlograms; version 1.6.9 used
library(ncf)              #for spline correlograms; version 1.2-5 used
library(spdep)            #for general correlograms; version 0.7-8 used
library(gstat)            #for variograms/kriging; version 1.1-6 used
library(terra)            #for raster covariate data; version 2.6-7 used
```


###############################################
# Correlograms
###############################################

```{r}
#load data
matrix = read.csv('~/Desktop/ECOL_620/Course_materials/Week5/Lab5/data_for_lab5/cactus_matrix.csv', header=T)
```

```{r}
#inspect
head(matrix)
```


```{r}
#plot
plot(matrix[,"y"] ~ matrix[,"x"],
     pch=21, cex=1.2,
     bg=gray.colors(12)[cut(matrix[,"Height"], breaks = 12)])
```


```{r}
#plot distribution of vegetation height
hist(matrix[,"Height"], xlab="Vegetation height (cm)",
     main="histogram of vegetation height")
```


```{r}
#calculate distance matrix
coords = cbind(matrix$x, matrix$y)
colnames(coords) = c("x", "y")
distmat = as.matrix(dist(coords))
```


```{r}
#inspect
dim(distmat)
max(distmat)
```


```{r}
#Maximum distance to consider in correlogram/variogram (~1/2 to 2/3 total dist)
maxdist = 2/3*max(distmat)
maxdist
```


#--------------------------------------#
# Correlogram with pgirmess
#--------------------------------------#

```{r}
correlog.pgirmess = pgirmess::correlog(coords, matrix$Height, method="Moran",
                            nbclass=14, alternative = "two.sided")
```


```{r}
#Moran and P values for each distance class
round(correlog.pgirmess,2)
```


```{r}
#plot
plot(correlog.pgirmess[,1], correlog.pgirmess[,2],
     xlab="Distance (m)", ylab="Moran's I", col = ifelse(correlog.pgirmess[,3] < 0.05,'red','black'), pch=19)
abline(h=0)
```


#--------------------------------------#
# Correlograms with ncf 
#--------------------------------------#

```{r}
#Correlogram with non-parameteric test of significance
correlog.ncf = ncf::correlog(x = matrix$x, y = matrix$y, z = matrix$Height,
                         increment=5, resamp=99)

#plot
plot(correlog.ncf)
abline(h=0)
```


```{r}
#Spline correlogram with 95% pointwise bootstrap confidence intervals
spline.corr = spline.correlog(x = matrix$x, y = matrix$y, z = matrix$Height,
                               xmax = maxdist, resamp=99, type="boot")

#plot
plot(spline.corr)
```


#############################################
# Variograms in gstat
#############################################

```{r}
gstat.veg = matrix
coordinates(gstat.veg) = ~x + y
```


```{r}
#Empirical semivariogram
emp.gstat = variogram(Height ~ 1, cutoff=maxdist, width=3, gstat.veg)
```


```{r}
#plot variogram
plot(emp.gstat)
```


#----------------------------------#
# Anisotropy check
#----------------------------------#

```{r}
#Directional variogram in gstat
emp4.gstat = variogram(Height ~ 1, cutoff=maxdist, alpha=c(0,45,90,135), gstat.veg)
```


```{r}
#plot directional variogram
plot(emp4.gstat)
```


#--------------------------------------------------#
# Model-based variograms with least squares in gstat
#--------------------------------------------------#

```{r}
#inspect models that gstat can fit
vgm()#fits more models than geoR
show.vgms()#plots examples of the various variograms
```


```{r}
#Spherical variogram
sph.gstat = fit.variogram(emp.gstat, vgm("Sph")) 
```


```{r}
#Exponential variogram
exp.gstat = fit.variogram(emp.gstat, vgm("Exp"))
```


```{r}
#inspect
exp.gstat
sph.gstat
```


```{r}
#plot
plot(emp.gstat, exp.gstat)
plot(emp.gstat, sph.gstat)
```


#################################################
# Kriging and interpolation in gstat
#################################################

```{r}
new.grid.1m.gstat = expand.grid(x=seq(0,max(matrix$x),.5), y=seq(0,max(matrix$y),.5)) #need labels for coords
```


```{r}
#convert to sp object
gridded(new.grid.1m.gstat) = ~x + y
```


```{r}
krig.gstat = krige(Height ~ 1, gstat.veg, new.grid.1m.gstat, model = sph.gstat)
```


```{r}
#Inverse distance weighting in gstat
idw.gstat = idw(Height ~ 1, gstat.veg, new.grid.1m.gstat) #idp-the power used; default power=2
```


```{r}
plot(krig.gstat)
```


```{r}
plot(idw.gstat)
```



